---
title: "Oct 1 Activities: M&Ms; Survey; Sampling"
date: 9-30-2025
---

All three class activities are on this .rmd notebook

Activity #1: Activity: M&Ms - Central Limit Theorem

Working in groups, each group will get a handful of candy from the bag, and then count the number of red and black candies. Log those in this Google Sheet 

https://docs.google.com/spreadsheets/d/1KOeoVTq2W7L2T9sWjwh5U_2f59GsTeAylVGD53rNWww/edit?usp=sharing

(give your group a nice name).

Take a look at the values that we all get - what do you notice about the patterns? This is an exercise about the central limit theorem. 

The central limit theorem states that the sampling distribution of the mean of any independent, random variable will be normal or nearly normal, if the sample size is large enough.

Then discuss and try to answer these questions:

--How close is your group's mean to the true proportion of 0.5?
--Why do different groups get different results?
--What happens to individual extreme results when we combine all groups?
--How does this relate to polls and surveys in real life?

Write your answer below. And then you may eat the M&Ms.

**A:**

Most group have a ratio 10% brown and 90% others. We have pretty similar numbers at 91%
M and M's were distributed differently to each group. 
This changes the average However, in this situation it didn't change much. 
Polls are based on area that are distributed in different ways. This can impact the result of certin polls, meaning distribution can change no matter where you go. In real life the sample size is not always a good repersentation of the entire population.

# Activity #2:

Belonging Survey

Working in pairs, take a look back at the main findings from UMD's Belonging Survey:

https://president.umd.edu/internal/belonging


Specifically, focus on the "Sample Description" section and the Appendix. Discuss the differences and between the population and the sample, and what the document says about how the survey attempted to deal with them. Are there any decisions described in the document that raise questions in your minds? Detail the differences you see and your understanding of the methods they employed (it's ok if you don't understand all of them), and add your questions. You should discuss this as a pair but submit individually.

**A:**


The sample description is does not seem repersentivte of the entire population. This is especially true because staff answered the majority of questions while students were under represented A responce rate of 23% is very encouraging but the fact that it is unequal is still alarming.  Also, it should be noted that women had a higher response rate then men. Religious affiliation seemed unbalanced as well
One must remember you can't ever have a perfect sample and  some of these details can still help you understand campus opinion You must take into account though that it's not a total representation of the entire population 




# TASK 3: SAMPLING

Based on the class sampling demo (available on Elms), sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

**A**
```{r}

library(tidyverse)

```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80

**A**
```{r}
set.seed(42)
younguns <- tibble(
  age = sample(18:80, 1000, replace = TRUE)
)
```

### Task 3: Sample 100 rows of data from the dataset
**A**
```{r}
sample_100 <-younguns |> 
  sample_n(100)


```


### Task 4: Calculate the mean age and standard deviation of the sample
**A**
```{r}
sample_100 |> summarize(mean = mean(age), median = median(age), sd = sd(age))

```


### Task 5: Calculate the mean age and standard deviation of the population
**A**
```{r}
younguns |> summarize(mean = mean(age), median = median(age), sd = sd(age))

```


### Task 6: Compare the sample and population statistics

The sample size size is fairly representative of  the whole, and this is evident given  the mean and slandered deviation 



Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

Both the mean and standard deviations are very similar. As long as you mention sample margin for error it is fair put the data in your article. This is true even when you run the data multiple times. 

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
